---
title: "release_date_from_album"
author: "Akshay Sundar"
date: "11/28/2018"
output: html_document
---

This code snippet is to extract the release date for each song from the Spotify API, using the album information.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(spotifyr)
library(tidyverse)
library(httr)
```


First step to import the song list - 

```{r}

one_million_df <- read_csv("~/Documents/R Programming/Group Assignment/Data/1mill_songs_db.csv")
#one million df has duplicates based on artist and song
one_million_df <- unique(one_million_df[c('artist','song')])

summary(one_million_df)
head(one_million_df)
```


Initialize the Spotify Client IDs -
```{r}

Sys.setenv(SPOTIFY_CLIENT_ID = 'a98864ad510b4af6851331638eec170f')
Sys.setenv(SPOTIFY_CLIENT_SECRET = '6445326414dd4bf381afbc779b182223')

access_token <- get_spotify_access_token()

```


Write function to pull in Album name and Release - 

Inputs : Track name and Artist name
Output : Vector with Album name and Release Date

```{r}
get_album_info <- function(song, artist) {
  song <- tolower(song)
  artist <- tolower(artist)
  track_name = gsub(' ','%20', gsub("[^[:alnum:][:space:]]",'',song))
  artist_name =gsub(' ','%20', gsub("[^[:alnum:][:space:]]",'',artist))
  api <- str_glue('https://api.spotify.com/v1/search/?q=track:{track_name}%20artist:{artist_name}&type=track&limit=1')
  access_token <- get_spotify_access_token()
  result <- RETRY('GET', url = api, query = list(access_token = access_token), quiet = TRUE, times = 1, pause_min = 1) %>% content 
    tryCatch({
      rel_date <- as.character(suppressWarnings(result$tracks$items[[1]]$album$release_date))
      rel_year <- ifelse(is.na(rel_date),rel_date,substr(rel_date,1,4))
      album_name <- as.character(suppressWarnings(result$tracks$items[[1]]$album$name))
      combined_response <- c(album_name,rel_year)
      return (combined_response)
      #return (rel_date)
    }, error=function(error_cond) {
      if (error_cond$message == 'subscript out of bounds') {
        return('ERROR subscript')
      }
    }
    , error=function(error_cond) {
      if (error_cond$message =='$ operator is invalid for atomic vectors') {
        access_token <- get_spotify_access_token()
        result2 <- RETRY('GET', url = api, query = list(access_token = access_token), quiet = TRUE, times = 1, pause_min = 1) %>% content 
        print(error_cond$message)
        print(api)
        rel_date2 <- as.character(suppressWarnings(result2$tracks$items[[1]]$album$release_date))
        rel_year2 <- ifelse(is.na(rel_date2),rel_date2,substr(rel_date2,1,4))
        album_name2 <- as.character(suppressWarnings(result$tracks$items[[1]]$album$name))
        combined_response2 <- c(album_name2,rel_year2)
        return (combined_response2)
        #return (rel_date2)
      }
      else {
        print(error_cond$message)
        return('ERROR mysterious')
      }
    }
  )
}

get_album_info("comfortably numb","pink floyd")
```

Defining a Function that takes in a dataframe with the Track and Artist, and returns a data frame with the Album name and Release Year

```{r}
get_album_vec <- function(dfin){
  artist <- as.vector(dfin[,1])
  track <- as.vector(dfin[,2])
  Album <- rep("Error",nrow(dfin))
  Release <- rep("Error",nrow(dfin))
  for (i in 1:nrow(dfin)) {
    temp <- get_album_info(track[i,],artist[i,])
    Album[i] <- temp[1]
    Release[i] <- temp[2]
  }
  return(data.frame(Album,Release))
}

#with timing 
get_album_vec_timed <- function(dfin){
  artist <- as.vector(dfin[,1])
  track <- as.vector(dfin[,2])
  Album <- rep("Error",nrow(dfin))
  Release <- rep("Error",nrow(dfin))
  for (i in 1:nrow(dfin)) {
    temp <- get_album_info(track[i,],artist[i,])
    Album[i] <- ifelse(is.null(temp[1]),"trigger",temp[1])
    Release[i] <- ifelse(is.null(temp[2]),"trigger",temp[2])
  }
  #Sys.sleep(0.3)
  return(data.frame(Album,Release))
}
```


Testing with a sample data frame -
```{r}

sample_size <- 10000

sample_n <- one_million_df[sample(nrow(one_million_df),sample_size),]

b <- get_album_vec_timed(sample_n)

sample_n$album <- b[,1]
sample_n$release <- b[,2]

saved_data <- sample_n
write_csv(saved_data,path = "album_info_test.csv")
```


Don't run this bit just yet-
```{r, eval=FALSE}

# trial_df['spotify_uri'] %>% filter(!str_detect(spotify_uri,'Error'))

#merge back to original df
trial_df2 <- try[!is.na(try$spotify_uri),]
complete_df <- merge(one_million_df, trial_df, by.x= c('artist', 'song'), by.y=c('artist', 'song'), all.x=TRUE)

#second trial
complete_df2 <- merge(complete_df, trial_df2, by.x= c('artist', 'song'), by.y=c('artist', 'song'), all.x=TRUE)

```



