---
title: "Clustering using Features Sample"
author: "Akshay Sundar"
date: "12/3/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

set.seed(333)

library(tidyverse)
library(FactoMineR)
library(cluster)
```

Loading Sample Dataset with Features
```{r}
features_sample <- read_csv("~/Documents/R Programming/Group Assignment/rubasic/rubasic/working_files/150k_sample.csv")
```

Studying Dataset -
```{r}
summary(features_sample)

#converting year to numeric
features_sample$year <- as.numeric(features_sample$album_year_4dgt)
features_sample <- features_sample %>% select(-album_year,-album_year_4dgt)
```

Removing NAs in pull and tracks with duration > 30 mins
```{r}
features_sample_proc <- na.omit(features_sample)

features_sample_proc <- features_sample_proc %>%
  filter(duration_ms<1800000)
```

Clustering through K-Means
```{r}
cluster_year_k <- function(objet,year_taken,clust_size){
  filter_obj <- na.omit(objet) %>% filter(year==(year_taken))
  restr_obj <- filter_obj %>% select(danceability, energy, key, loudness, mode, speechiness, acousticness, instrumentalness, liveness, valence, tempo, duration_ms, time_signature, year)
  clust_features <- kmeans(restr_obj,centers=clust_size,iter.max = 1000, nstart = 100)
  filter_obj$k_cluster <- paste0(year_taken,"_",clust_features$cluster)
  return(list(filter_obj,clust_features))
}
```

Clustering through HCPC
```{r}
cluster_year_hcpc <- function(objet,year_taken,clust_num){
  filter_obj <- na.omit(objet) %>% filter(year==(year_taken))
  restr_obj <- filter_obj %>% select(danceability, energy, key, loudness, mode, speechiness, acousticness, instrumentalness, liveness, valence, tempo, duration_ms, time_signature, year)
  clust_features <- HCPC(restr_obj, nb.clust = clust_num, graph = F)
  filter_obj$hcpc_cluster <- paste0(year_taken,"_",clust_features$data.clust$clust)
  return(list(filter_obj,clust_features))
}
```


Clustering through HCPC with pre PCA (2 principal dimensions)
```{r}
cluster_year_hcpc_pca <- function(objet,year_taken,clust_num){
  filter_obj <- na.omit(objet) %>% filter(year==(year_taken))
  restr_obj <- filter_obj %>% select(danceability, energy, key, loudness, mode, speechiness, acousticness, instrumentalness, liveness, valence, tempo, duration_ms, time_signature, year)
  pca_res <- PCA(restr_obj,ncp=2,graph = F)
  clust_features <- HCPC(pca_res, nb.clust = clust_num, graph = F)
  filter_obj$dim_1 <- pca_res$ind$coord[,1]
  filter_obj$dim_2 <- pca_res$ind$coord[,2]
  filter_obj$hcpc_pca_cluster <- paste0(year_taken, "_", clust_features$data.clust$clust)
  return(list(filter_obj,clust_features))
}
```

K means clusters database
```{r}
min_year <- min(features_sample$year)
max_year <- max(features_sample$year)
features_clustered_k <- data.frame()
model_k_clusters <- list()

for (year in min_year:max_year) {
    temp <- cluster_year_k(features_sample_proc,year,3)
    features_clustered_k <- rbind(features_clustered_k,as.data.frame(temp[1]))
    model_k_clusters <- append(model_k_clusters,temp[2])
}
```

HCPC clusters database
```{r}
features_clustered_hcpc <- data.frame()
model_hcpc_clusters <- list()
    
for (year in min_year:max_year) {
    temp <- cluster_year_hcpc(features_sample_proc,year,3)
    features_clustered_hcpc <- rbind(features_clustered_hcpc,as.data.frame(temp[1]))
    model_hcpc_clusters <- append(model_hcpc_clusters,temp[2])
}
```

HCPC with pre PCA clusters database
```{r}
features_clustered_hcpc_pca <- data.frame()
model_hcpc_pca_clusters <- list()
    
for (year in min_year:max_year) {
    temp <- cluster_year_hcpc_pca(features_sample_proc,year,3)
    features_clustered_hcpc_pca <- rbind(features_clustered_hcpc_pca,as.data.frame(temp[1]))
    model_hcpc_pca_clusters <- append(model_hcpc_pca_clusters,temp[2])
}
```

Combination of Cluster databases
```{r}
combined_clusters <- features_clustered_hcpc_pca
combined_clusters$hcpc_cluster <- features_clustered_hcpc$hcpc_cluster
combined_clusters$k_cluster <- features_clustered_k$k_cluster

#Output
write_csv(combined_clusters,'~/Documents/R Programming/Group Assignment/rubasic/rubasic/working_files/150K_samples_clustered.csv')
```


Garbage testing from here on -
```{r}

temp <- cluster_year_hcpc_pca(features_sample_proc,2008,3)
temp1 <- as.data.frame(temp[1])

nums <- unlist(lapply(temp, is.numeric))
nums["cluster"] <- TRUE
temp1 <- temp[,nums] %>% select(-X1,-X) %>% group_by(cluster) %>% summarise_all(funs(mean))
temp[,nums] %>% select(-X1,-X) %>% group_by(cluster) %>% summarise(n=n())
temp1

ggplot(data=temp) + geom_point(aes(x=dim_1,y=dim_2,colour=cluster))
```

```{r}
nums <- unlist(lapply(features_sample, is.numeric))
temp1 <- features_sample[,nums] %>% select(-X1,-X) %>% filter(year==2010)
clust_features <- HCPC(temp1, nb.clust=5)
clust_features$call$t
clust_features$data.clust

```

