---
title: "descr_clara"
author: "Clara Dionet"
date: '2018-12-12'
output: html_document
---

Motivations & Learning

In order to visualize and compare the evolution of top and average song attributes (such as "valence", "danceability", etc.), we decided to implement the attributes_time function. 

This function plots overlapping variables (chosen by the user) of 2 different data sets as a function of the "year" column. In our case these two dataframes correspond to top and average songs ("Billboard" and "Non Billboard" respectively).

The choice to take as arguments 2 dataframes instead of 1 (with an additional column specifying the category of each dataset) was to avoid unnecessary formatting in case of very different datasets. If the user only needs to plot one overlapping feature in his 2 datasets, he does not need to match every column name to each other before applying the rbind function in order to store them as one dataframe. 

This was appropriate in our case as the billboard data was collected from a different source than the one for average songs, but both contained music features from the Spotify API.

The input arguments are as follows:

First dataset characteristics:
- df1: a data.frame
- df1_title: title of your first df (ex: topsongs)
- df1_year_col: a number (column number of the "year" column)

Second dataset characteristics:
- df2: a data.frame
- df2_title: title of your second df (ex: averagesongs)
- df2_year_col: a number (column number of the "year" column)

Plot characteristics:
- attributes: a vector containing strings, the column names to be plotted (ex: c("energy","danceability"))
- boxplot: a boolean, specifies the type of plot (TRUE = plot as boxplots, FALSE = plot as averages)
- timerange: a 2x1 vector, this defines xlims, years to be plotted over (ex: c(1990, 2010))
- title_vector: a string vector with the df_title names (the strings have to match!). This defines what will be plotted, for example, c(df1_title) will plot only the first dataframe, and c(df1_title, df2_title) will plot both. Although this is a redundant input parameter, it was created to facilitate the synchronization with the shiny app (the user can choose from the "Billboard" or "Non Billboard" checkboxes).

Example: 
Load data & define inputs:
```{r}
library(musictasteR)
data("billboard_music_dataframe")
data("averagesongs")

topsongs <- averagesongs
```

```{r}
head(averagesongs)
head(topsongs)
```
The variable "year" is in column 1 for topsongs and column 4 for average songs.

Comparing averages for Billboard / Non Billboard data:
```{r}
attributes <- c("danceability" ,"energy" ,"mode", "speechiness","acousticness", "instrumentalness" ,"liveness","valence" ,"tempo")
boxplot <- FALSE
timerange <- c(1960,2015)
title_vector <- c("Top Songs","Average Songs") # plot both

attributes_time(topsongs, "Top Songs", 1, averagesongs, "Average Songs", 4, attributes, boxplot, timerange, title_vector)
```

Plot only billboard data:
```{r}
title_vector <- "Average Songs"
attributes_time(topsongs, "Top Songs", 1, averagesongs, "Average Songs", 4, attributes, boxplot, timerange, title_vector)

```

Plot as boxplots:
```{r}
attributes <- c("danceability" ,"energy")
boxplot <- TRUE
attributes_time(topsongs, "Top Songs", 1, averagesongs, "Average Songs", 4, attributes, boxplot, timerange, title_vector)
```

